{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2019-02-08 papers/methods/figures/tikz.tex
Updating figure again
Removing p

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \hat{\mathcal{C}}_\eta^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[det, above=of obssum, xshift=1.2cm] (mb) {$m_B$};
	\node[latent, left=of mb, xshift=0.1cm] (x1c) {$x_1, c$};
	\node[det, above=of zobs] (z) {$z$};
	\node[det, above=of mobs, yshift=-0.5cm] (m) {$m$};
	\node[det, above=of pobs, yshift=0.5cm] (p) {$p$};	
	\node[det, above=of z] (mu) {$\mu$};
	\node[latent, above=of p, yshift=-0.5cm] (MB) {$M_B$};
	% Survey nodes
	\node[latent, above=of x1c, yshift=0.9cm, xshift=-0.65cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };

	\node[latent, right=0.2cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
	% Global nodes
	\node[latent, above=2.1cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=1.8cm of MB, xshift=0cm, align=center]  (MBpop) {$\langle M_B \rangle,$ \\ $\sigma_{M_B}$};
	\node[latent, right=1.43cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
	\node[latent, right=0.1cm of dz, align=center] (ab) {$\alpha, \beta, \delta(0)$, \\$\delta(\infty)$};
	% Connect the nodes
	\edge {mb, x1c, dz, kappa} {obssum};
	\edge {z} {zobs};
	\edge {m} {mobs};
	\edge {p} {pobs};
	\edge {z,cosmo} {mu};
	\edge {MBpop} {MB};	
	\edge {MB, mu, ab, m, p, x1c} {mb}
	\edge {pop1, z} {x1c};

	% Plates
	\plated[thick] {sn} {(mb)(obssum)(zobs)(mobs)(pobs)(x1c)(z)(mu)(MB)(m)(p)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)(kappa)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

>>>>

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \hat{\mathcal{C}}_\eta^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.9cm of obssum]  (mobs) {$\hat{{\rm mass}}$};

	% Latent nodes
	\node[det, above=of obssum, xshift=1.5cm] (mb) {$m_B$};
	\node[latent, left=of mb, xshift=-0.2cm] (x1c) {$x_1, c$};
	\node[det, above=of zobs] (z) {$z$};
	\node[det, above=of mobs, yshift=-0.5cm] (m) {${\rm mass}$};

	\node[det, above=of z] (mu) {$\mu$};
	\node[latent, above=of m, xshift=0.8cm, yshift=-0.5cm] (MB) {$M_B$};
	% Survey nodes
	\node[latent, above=of x1c, yshift=0.9cm, xshift=-0.65cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };

	\node[latent, right=0.2cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
	% Global nodes
	\node[latent, above=2.1cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=2.5cm of MB, xshift=0cm, align=center]  (MBpop) {$\langle M_B \rangle,$ \\ $\sigma_{M_B}$};
	\node[latent, right=1.43cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
	\node[latent, right=0.1cm of dz, align=center] (ab) {$\alpha, \beta,$\\ $\delta(0),\delta(\infty)$};
	% Connect the nodes
	\edge {mb, x1c, dz, kappa} {obssum};
	\edge {z} {zobs};
	\edge {m} {mobs};
	\edge {z,cosmo} {mu};
	\edge {MBpop} {MB};	
	\edge {MB, mu, ab, m, x1c} {mb}
	\edge {pop1, z} {x1c};

	% Plates
	\plated[thick] {sn} {(mb)(obssum)(zobs)(mobs)(x1c)(z)(mu)(MB)(m)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)(kappa)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

### 2019-02-08 papers/methods/paper.tex
Removing p

\subsubsection{Observables}
Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$. Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses. We also have a probability of each supernova being a type Ia, $\hat{p}$. Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

@@

\begin{deluxetable}{ll}
	\renewcommand{\arraystretch}{0.8}
	\tablecolumns{2}
	\tablewidth{0.99\columnwidth}
	\tablecaption{Model Parameters}
	\tablehead { \colhead{Parameter} & \colhead{Description} }
	\startdata	
	&\textbf{Global Parameters} \\
	$\Omega_m$  & Matter density  \\
	$w$  &  Dark energy equation of state  \\
	$\alpha$ & Stretch standardisation   \\
	$\beta$         &  color standardisation   \\
	$\delta(0)$ & Scale of the mass-magnitude correction\\
	$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
	$\delta\mathcal{Z}_i$ & Systematics scale\\
	$\langle M_B \rangle$ & Mean absolute magnitude \\
	\\
	& \textbf{Survey Parameters} \\
	$\delta S$ & Selection effect deviation \\
	$\langle x_1^i \rangle$ & Mean stretch nodes\\
	$\langle c^i \rangle$ & Mean color nodes\\
	$\alpha_c$ & Skewness of color population \\
	$\sigma_{M_B}$ & Population magnitude scatter \\
	$\sigma_{x_1}$ & Population stretch scatter \\
	$\sigma_{c}$ & Population color scatter \\
	$\kappa_{0}$ & Extra color dispersion\\
	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
	\\
	& \textbf{Supernova Parameters} \\
	$m_B$ & True flux\\
	$x_1$ & True stretch \\
	$c$ & True color \\
	$z$ & True redshift \\	
	$M_B$ & Derived absolute magnitude \\
	$\mu$ & Derived distance modulus \\
	\\
	& \textbf{Input Data}\tablenotemark{a} \\
	$\hat{m}_B$ & Measured flux\\
	$\hat{x}_1$ & Measured stretch \\
	$\hat{c}$ & Measured color \\
	$C$ & Covariance on flux, stretch and color \\
	$\hat{z}$ & Observed redshift \\
	$\hat{m}$ & Determined mass probability \\
	$\hat{p}$ & Determined Ia probability \\
	\enddata
	\tablenotetext{a}{Not model parameters but shown for completeness.}
\label{tab:param_summary}
\end{deluxetable}

@@

 \begin{figure}
 	\begin{center}
 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
 	\end{center}
 	\caption{Probabilistic graphical model for our likelihood \added{without selection effects}. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables. The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it. Variables that appear outside both boxes represent top level model parameters. \added{We note that we have shown the model to have latent variables $\lbrace M_B,\,x_1\,c\rbrace$, which uniquely determines $m_B$, given $\mu$ and other parameters. Thus, the two nodes $M_B$ and $m_B$, make up a single layer in our hierarchy, not two layers. In the code implementation, $m_B$ is more efficiently parametrised instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other. We write out ${\rm mass}$ instead of $m$ to reduce possible confusion with magnitude in the diagram. Finally, as we take redshift measurement $\hat{z}$, mass probability $\hat{{\rm mass}}$ and classification probability $\hat{p}$ as exact they are not conditioned on underlying distributions and are top level parameters. We add the deterministic $m$ and $p$ nodes in our diagram for clarity, however as they have a one-to-one mapping onto $\hat{{\rm mass}}$ and $\hat{p}$ respectively we simply use the observed values in the mathematics of the rest of the paper.}}
 	\label{fig:pgm}
 \end{figure}

@@

For future use with photometrically classified surveys, we include a simplistic outlier population model. We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works. However, with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.  We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section. For the spectroscopic survey, we set this to unity. For the photometric proof-of-concept we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates will be left for future improvements during the analysis of the DES photometric sample.

@@

There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}. The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step. The scale of this step function varies from analysis to analysis, and is treated as a fit parameter. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}
where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift. We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.

@@

\begin{align}
&P({\rm data}|\theta) = \notag \\
&\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z) \notag \\
& P(\mu |z, \Omega_m, w) 
\end{align}

@@

\subsection{Supernova Selection Effects}
\label{app:selection2}
\added{
To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters. We thus continue from equation \eqref{eq:latent} and write out denominator $d$:
\begin{equation}
d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL.
\end{equation}
Here we can now move from the generic labels $D$ and $L$ to something specific to our model. In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters. That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$. Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity. If we did, we would simply find them integrated to unity. Note that this assumption does not capture biases engendered by
Poisson noise fluctuations.  
We advocate that future analyses with
higher statistical precision do use a precisely determined $P(S|D)$.
Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
\begin{equation}
d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
\end{equation}
Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$. Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{equation}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{equation}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{equation}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{equation}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

>>>>

\subsubsection{Observables}
Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$. Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses. Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

@@

\begin{deluxetable}{ll}
	\renewcommand{\arraystretch}{0.8}
	\tablecolumns{2}
	\tablewidth{0.99\columnwidth}
	\tablecaption{Model Parameters}
	\tablehead { \colhead{Parameter} & \colhead{Description} }
	\startdata	
	&\textbf{Global Parameters} \\
	$\Omega_m$  & Matter density  \\
	$w$  &  Dark energy equation of state  \\
	$\alpha$ & Stretch standardisation   \\
	$\beta$         &  color standardisation   \\
	$\delta(0)$ & Scale of the mass-magnitude correction\\
	$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
	$\delta\mathcal{Z}_i$ & Systematics scale\\
	$\langle M_B \rangle$ & Mean absolute magnitude \\
	\\
	& \textbf{Survey Parameters} \\
	$\delta S$ & Selection effect deviation \\
	$\langle x_1^i \rangle$ & Mean stretch nodes\\
	$\langle c^i \rangle$ & Mean color nodes\\
	$\alpha_c$ & Skewness of color population \\
	$\sigma_{M_B}$ & Population magnitude scatter \\
	$\sigma_{x_1}$ & Population stretch scatter \\
	$\sigma_{c}$ & Population color scatter \\
	$\kappa_{0}$ & Extra color dispersion\\
	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
	\\
	& \textbf{Supernova Parameters} \\
	$m_B$ & True flux\\
	$x_1$ & True stretch \\
	$c$ & True color \\
	$z$ & True redshift \\	
	$M_B$ & Derived absolute magnitude \\
	$\mu$ & Derived distance modulus \\
	\\
	& \textbf{Input Data}\tablenotemark{a} \\
	$\hat{m}_B$ & Measured flux\\
	$\hat{x}_1$ & Measured stretch \\
	$\hat{c}$ & Measured color \\
	$C$ & Covariance on flux, stretch and color \\
	$\hat{z}$ & Observed redshift \\
	$\hat{m}$ & Observed mass probability \\
	\enddata
	\tablenotetext{a}{Not model parameters but shown for completeness.}
\label{tab:param_summary}
\end{deluxetable}

@@

 \begin{figure}
 	\begin{center}
 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
 	\end{center}
 	\caption{Probabilistic graphical model for our likelihood \added{without selection effects}. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables. The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it. Variables that appear outside both boxes represent top level model parameters. \added{We note that we have shown the model to have latent variables $\lbrace M_B,\,x_1\,c\rbrace$, which uniquely determines $m_B$, given $\mu$ and other parameters. Thus, the two nodes $M_B$ and $m_B$, make up a single layer in our hierarchy, not two layers. In the code implementation, $m_B$ is more efficiently parametrised instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other. We write out ${\rm mass}$ instead of $m$ to reduce possible confusion with magnitude in the diagram. Finally, as we take redshift measurement $\hat{z}$ and mass probability $\hat{{\rm mass}}$ as exact they are not conditioned on underlying distributions and are top level parameters.}}
 	\label{fig:pgm}
 \end{figure}

@@

\added{For the spectroscopic paper we do not consider outlier populations, however we ensure that our model has flexibility for such populations for future use with photometrically classified surveys. We thus include a simplistic outlier population model.} We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model, where an additional observable of the SN~Ia probability would be needed in order to inform the weights of the mixture model. For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works. However, with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.  We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.  We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude. We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$ in our tests. The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$. For the spectroscopic survey, we set this to unity, and thus it is not included in Figure \ref{fig:pgm} or Table \ref{tab:param_summary}. For the photometric proof-of-concept we provide an accurate probability estimate. Further investigation on the effect of inaccurate estimates will be left for future improvements during the analysis of the DES photometric sample.

@@

There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}. The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step. The scale of this step function varies from analysis to analysis, and is treated as a fit parameter. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}
where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift. We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$. \added{Finally, we assume that the observed mass probability $\hat{m}$ supplied to the model is perfectly determined, and thus set $P(\hat{m}|m) = \delta(\hat{m} - m)$.}

@@

\begin{align}
&P({\rm data}|\theta) \propto \notag \\
&\int\, d\mu\, dm_B\, dx_1\, dc\, dz, dm \  P(\hat{z}|z) P(\hat{m}|m)  \notag \\
& P(\hat{m}_B, \hat{x}_1, \hat{c}, \hat{\mathcal{C}}_\eta^{-1} | m_B, x_1, c, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) \notag \\
& P(x_1, c |  \langle x_1^i \rangle, \langle c^i \rangle, \sigma_{x_1}, \sigma_c, z) \notag\\
& P(m_B | \langle M_B \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z, m, p) \notag \\
& P(\mu |z, \Omega_m, w).
\end{align}

@@

\subsection{Supernova Selection Effects}
\label{app:selection2}
\added{
To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters. We thus continue from equation \eqref{eq:latent} and write out denominator $d$:
\begin{equation}
d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL.
\end{equation}
Here we can now move from the generic labels $D$ and $L$ to something specific to our model. In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m} \rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters. That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$. Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity. If we did, we would simply find them integrated to unity. Note that this assumption does not capture biases engendered by
Poisson noise fluctuations.  
We advocate that future analyses with
higher statistical precision do use a precisely determined $P(S|D)$.
Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
\begin{equation}
d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
\end{equation}
Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$. Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{equation}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{equation}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{equation}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{equation}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.
